---
title: "exercise-05"
author: "Riley N Derby"
date: "2024-02-28"
output: html_document
---

step 1: lets read in the file 
```{r}
library(tidyverse)
library(dplyr)

f <- "https://raw.githubusercontent.com/difiore/ada-2024-datasets/main/IMDB-movies.csv"
d <- read_csv(f, col_names = TRUE)  # creates a 'tibble'
head(d)
glimpse(d)
summary(d)
```

step 2: use 1 line to filter dataset to 
1.movies from 1920-1979 
2.movies that are between 1-3 hours long
3. make a new column codes startYear into new variable called decade ie 20s, 30s 40s

```{r}
d <- d %>%
  filter(startYear >= 1920 & startYear <= 1979) %>%
  filter(runtimeMinutes >= 60 & runtimeMinutes <= 180) %>%
  mutate(decade = case_when(
    startYear >= 1920 & startYear < 1930 ~ "1920s",
    startYear >= 1930 & startYear < 1940 ~ "1930s",
    startYear >= 1940 & startYear < 1950 ~ "1940s",
    startYear >= 1950 & startYear < 1960 ~ "1950s",
    startYear >= 1960 & startYear < 1970 ~ "1960s",
    startYear >= 1970 & startYear < 1980 ~ "1970s",
    TRUE ~ NA_character_))
```

i have 5651 remaining movies with a new decade column so this step is complete


lets move to step 3: use ggplot to plot histogram of distribution of runtimeminutes for each decade

```{r}
histogram <- ggplot(d, aes(x = runtimeMinutes)) +
  geom_histogram(binwidth = 10, fill = "blue", color = "red") +
  facet_wrap(~ decade, scales = "free") +
  theme_classic()
histogram
```


created histograms for each decade

now lets move on 

step 4: Use a one-line statement to calculate the population mean and population standard deviation in runtimeMinutes for each decade and save the results in a new dataframe called results.
```{r}
results <- d %>%
  group_by(decade) %>%
  summarise(mean = mean(runtimeMinutes, na.rm = T),
            sd = sd(runtimeMinutes, na.rm = T))
  
head(results)
```

now that i have my results dataframe we can move to next step

step 5: Draw a single sample of 100 movies, without replacement, from each decade and calculate the single sample mean and single sample standard deviation in runtimeMinutes for each decades. Recall that your single sample mean for each decade is an estimate of the population mean for each decade.

```{r}

# Sample size
n <- 100

# Function to draw a single sample grouped by decade and calculate mean and standard deviation
draw_sample <- group_by(d, decade) %>%
  sample_n(n, replace = F) %>%
  summarise(samp_mean = mean(runtimeMinutes, na.rm = T),
            samp_sd = sd(runtimeMinutes, na.rm = T))

draw_sample
```


step 6: Calculate for each decade the standard error around your estimate of the population mean runtimeMinutes based on the standard deviation and sample size (n=100 movies) of your single sample

```{r}
se_samp <- draw_sample$samp_sd/sqrt(n)
se_samp
```

combine se_samp with draw_sample dataset by rerunning and adding it to summarize function
```{r}
draw_sample <- group_by(d, decade) %>%
  sample_n(n, replace = F) %>%
  summarise(samp_mean = mean(runtimeMinutes, na.rm = T),
            samp_sd = sd(runtimeMinutes, na.rm = T),
            se_samp = samp_sd/sqrt(n))

draw_sample
```

step 7: Compare these estimates to the actual population mean runtimeMinutes for each decade and to the calculated SE in the population mean for samples of size 100 based on the population standard deviation for each decade.
```{r}
#recall actual datset name
results

#call up sample dataset
draw_sample


#lets join them by decade
compare <- left_join(draw_sample, results, by = "decade")
compare
```

step 8: Generate a sampling distribution of mean runtimeMinutes for each decade by [a] drawing 1000 random samples of 100 movies from each decade, without replacement, and, for each sample,  [b] calculating the mean runtimeMinutes and the standard deviation in runtimeMinutes for each decade. Use either a standard for( ){ } loop, the do(reps) * formulation from {mosaic}, the rerun() function from {purrr}, or the rep_sample_n() workflow from {infer} to generate your these sampling distributions (see Module 16).

```{r}
library(mosaic)
reps = 1000

sampling_dist <- do(reps) * sample_n(group_by(d, decade), n, replace = F) %>%
  group_by(decade) %>%
  summarise(samp_dist_mean = mean(~runtimeMinutes, na.rm = T),
            samp_dist_sd = sd(~runtimeMinutes, na.rm = T))
  

head(sampling_dist)
```


step 9: Then, calculate the mean and the standard deviation of the sampling distribution of sample means for each decade (the former should be a very good estimate of the population mean, while the latter is another estimate of the standard error in our estimate of the population mean for a particular sample size) and plot a histogram of the sampling distribution for each decade. What shape does it have?

```{r}
sample_results <- sampling_dist %>%
  group_by(decade) %>%
  summarise(mean_samp_dist = mean(samp_dist_mean, na.rm = T),
            sd_samp_dist = sd(samp_dist_mean, na.rm = T),
            se_samp_dist = sd_samp_dist/sqrt(reps)) 
sample_results
```


histogram of each decade-- all turn out to have normal distribution
```{r}
samp_dist_plot <- ggplot(data = sampling_dist, aes(x = samp_dist_mean)) + 
  geom_histogram() + facet_wrap(~decade)
samp_dist_plot
```

step 10: Finally, compare the standard error in runtimeMinutes for samples of size 100 from each decade [1] as estimated from your first sample of 100 movies, [2] as calculated from the known population standard deviations for each decade, and [3] as estimated from the sampling distribution of sample means for each decade.

```{r}
final_compare <- left_join(compare, sample_results)
final_compare
```


###########################################################################


alright lets do challenge 2


step 1: read in zombies csv
```{r}
g <- "https://raw.githubusercontent.com/difiore/ada-2024-datasets/main/zombies.csv"
z <- read_csv(g, col_names = TRUE)  # creates a 'tibble'
head(z)
glimpse(z)
summary(z)
```

step 2: Calculate the population mean and standard deviation for each quantitative random variable in the dataset (height, weight, age, number of zombies killed, and years of education).
```{r}
##calculate the means

pop_mean <- z %>%
  summarise(mean(height), mean(weight), mean(age), mean(zombies_killed), mean(years_of_education))
pop_mean


#calculate popsd
height_sd <- sd(z$height)*(sqrt((length(z$height)-1)/length(z$height)))
height_sd #4.30797

weight_sd <- sd(z$weight)*(sqrt((length(z$weight)-1)/length(z$weight)))
weight_sd #18.39186

age_sd <- sd(z$age)*(sqrt((length(z$age)-1)/length(z$age)))
age_sd #2.963583

zk_sd <- sd(z$zombies_killed)*(sqrt((length(z$zombies_killed)-1)/length(z$zombies_killed)))
zk_sd #1.747551

ye_sd <- sd(z$years_of_education)*(sqrt((length(z$years_of_education)-1)/length(z$years_of_education)))
ye_sd #1.675704

```

step 3: Use {ggplot} and make boxplots of each of these variables by gender

```{r}
library(ggplot2)

gender_height <- ggplot(data = z, aes(x = gender, y = height)) +
  geom_boxplot(na.rm = T)
gender_height

gender_weight <- ggplot(data = z, aes(x = gender, y = weight)) +
  geom_boxplot(na.rm = T)
gender_weight

gender_age <- ggplot(data = z, aes(x = gender, y = age)) +
  geom_boxplot(na.rm = T)
gender_age

gender_zk <- ggplot(data = z, aes(x = gender, y = zombies_killed)) +
  geom_boxplot(na.rm = T)
gender_zk

gender_ye <- ggplot(data = z, aes(x = gender, y = years_of_education)) +
  geom_boxplot(na.rm = T)
gender_ye


```

step 4: Use {ggplot} and make scatterplots of height and weight in relation to age (i.e., use age as the x variable), using different colored points for males versus females. Do these variables seem to be related? In what way?

```{r}
age_height <- ggplot(data = z, aes(x = age, y = height, color = factor(gender))) +
  geom_point(na.rm = T)
age_height


age_weight <- ggplot(data = z, aes(x = age, y = weight, color = factor(gender))) +
  geom_point(na.rm = T)
age_weight
```

these variables seem positively realted i.e as age increases, height and weight increases for both sexes. Additionally, males seem to be generally taller and heavier than females 



step 5:Using histograms and Q-Q plots, check whether each of the quantitative variables seem to be drawn from a normal distribution. Which seem to be and which do not? 

```{r}
## qq plots
height_q <- qqnorm(z$height, pch = 1, frame = FALSE)
height_q

weight_q <- qqnorm(z$weight, pch = 1, frame = FALSE)
weight_q

age_q <- qqnorm(z$age, pch = 1, frame = FALSE)
age_q

zk_q <- qqnorm(z$zombies_killed, pch = 1, frame = FALSE)
zk_q

ye_q <- qqnorm(z$years_of_education, pch = 1, frame = FALSE)
ye_q

##histograms
height_hist <- ggplot(data = z, aes(x  = height)) +
  geom_histogram(bins = 30)
height_hist

weight_hist <- ggplot(data = z, aes(x  = weight)) +
  geom_histogram(bins = 30) 
weight_hist

age_hist <- ggplot(data = z, aes(x  = age)) + 
  geom_histogram(bins = 30) 
age_hist

zk_hist <- ggplot(data = z, aes(x  = zombies_killed)) +
  geom_histogram(bins = 30)
zk_hist

ye_hist <- ggplot(data = z, aes(x  = years_of_education)) +
  geom_histogram(bins = 30)
ye_hist


```

height, weight, and age are normally distributed... whereas zombies killed and years of education have a left-skew


step 6: Now use the sample_n() or slice_sample() function from {dplyr} to sample ONE subset of 50 zombie apocalypse survivors (without replacement) from this population and calculate the mean and sample standard deviation for each variable. Also estimate the standard error for each variable based on this one sample and use that to construct a theoretical 95% confidence interval for each mean. You can use either the standard normal or a Student’s t distribution to derive the critical values needed to calculate the lower and upper limits of the CI.

```{r}

#we need to create a loop that cycles through the variables of interest so im gonna make an empty tibble with just those variables

m <- 50
samp1 <- sample_n(z, size = m, replace = F)

y <- select(samp1, height, weight, age, zombies_killed, years_of_education)

samp50 <- tibble(variable = character(), samp_mean = numeric(), samp_sd = numeric(), samp_se = numeric(), lower_ci = numeric(), upper_ci = numeric())


#okay so we have our isolated dataset called y and an empty tibble samp50 to put the information from the loop into
for (i in seq_along(y)) {
  samp_mean <- mean(y[[i]])
  samp_sd <- sd(y[[i]])
  samp_se <- samp_sd/sqrt(m)
  lower_ci <- samp_mean + qnorm(0.025 / 2) 
  upper_ci <-  samp_mean - qnorm(0.025 / 2)
  samp50 <- add_row(samp50, variable = (colnames(y)[i]), samp_mean = samp_mean, samp_sd = samp_sd, samp_se = samp_se, lower_ci = lower_ci, upper_ci = upper_ci)
}

samp50
```

step 7: Then draw another 199 random samples of 50 zombie apocalypse survivors out of the population and calculate the mean for each of the these samples. Together with the first sample you drew out, you now have a set of 200 means for each variable (each of which is based on 50 observations), which constitutes a sampling distribution for each variable. What are the means and standard deviations of the sampling distribution for each variable? How do the standard deviations of the sampling distribution for each variable compare to the standard errors estimated from your first sample of size 50?

```{r}
library(mosaic)

reps <- 199

#create another tibble for 199 new samples

samp_199 <- tibble(
  z_seq = c(1:reps),
  height_mean = do(reps) * mean(~height, data = sample_n(z, size = m, replace = F)),
  weight_mean = do(reps) * mean(~weight, data = sample_n(z, size = m, replace = F)),
  age_mean = do(reps) * mean(~age, data = sample_n(z, size = m, replace = F)),
  zk_mean = do(reps) * mean(~zombies_killed, data = sample_n(z, size = m, replace = F)),
  ye_mean = do(reps) * mean(~years_of_education, data = sample_n(z, size = m, replace = F))
  )

samp_199

head(samp_199)
```




cant figure out how to combine the two df together...
```{r}
## we need to combine samp199 with samp50 to have our distribution of 200

## just do it by hand

samp50_wider <- tibble(
  z_seq = 200,
  height_mean = list(mean = 68.35810),
  weight_mean = list(mean = 145.28854),
  age_mean = list(mean = 20.54319),
  zk_mean = list(mean = 3.14),
  ye_mean = list(mean = 3.08)
)

# Add the new row to samp_199
samp_200<- bind_rows(samp_199, samp50_wider)

```


going to just rock it by hand and combine
```{r}

head(samp50)


height_mean_dist <- (mean(~mean, data = samp_199$height_mean) +68.35810)/2
height_mean_dist

weight_mean_dist <- (mean(~mean, data = samp_199$weight_mean) +145.28854)/2
weight_mean_dist


age_mean_dist <- (mean(~mean, data = samp_199$age_mean) +20.54319)/2
age_mean_dist

zk_mean_dist <- (mean(~mean, data = samp_199$zk_mean) +3.14)/2
zk_mean_dist


ye_mean_dist <- (mean(~mean, data = samp_199$ye_mean) +3.08)/2
ye_mean_dist

```

standard deviations
```{r}
h_meanSD <- sd(~mean, data = samp_199$height_mean) 
h_meanSD
w_meanSD <- sd(~mean, data = samp_199$weight_mean) 
w_meanSD
age_meanSD <- sd(~mean, data = samp_199$age_mean) 
age_meanSD
zk_meanSD <- sd(~mean, data = samp_199$zk_mean) 
zk_meanSD
ye_meanSD <- sd(~mean, data = samp_199$ye_mean)	
ye_meanSD
```

the standard deviations are now very similar to that of the se from first sample




step 8: Plot the sampling distributions for each variable mean. What do they look like? Are they normally distributed? What about for those variables that you concluded were not originally drawn from a normal distribution?

```{r}
# Define the names of the variables
variable_names <- c("height_mean", "weight_mean", "age_mean", "zk_mean", "ye_mean")

# Create histograms for each variable
for (var in variable_names) {
  hist(unlist(samp_199[[var]]), main = var, xlab = var, col = "skyblue", border = "white")
}

```

they look pretty normal for the most part... higher reps would make it better


Step 9
Construct a 95% confidence interval for each mean directly from the sampling distribution of sample means using the central 95% that distribution (i.e., by setting the lower and upper CI bounds to 2.5% and 97.5% of the way through that distribution).


lets change my samp_199 to numeric instead of df
```{r}
# Function to extract numeric value from nested data frame
extract_numeric <- function(x) {
  as.numeric(unlist(x$mean))
}

# Apply the function to each column
samp_199_num <- samp_199 %>%
  mutate(
    height_mean = extract_numeric(height_mean),
    weight_mean = extract_numeric(weight_mean),
    age_mean = extract_numeric(age_mean),
    zk_mean = extract_numeric(zk_mean),
    ye_mean = extract_numeric(ye_mean)
  )

# View the updated dataframe
print(samp_199_num)

```



```{r}
lower <- quantile(samp_199_num$height_mean, 0.025)
upper <- quantile(samp_199_num$height_mean, 0.975)
(ci_height <- c(lower, upper))

lowerw <- quantile(samp_199_num$weight_mean, 0.025)
upperw <- quantile(samp_199_num$weight_mean, 0.975)
(ci_weight <- c(lowerw, upperw))

lowera <- quantile(samp_199_num$age_mean, 0.025)
uppera <- quantile(samp_199_num$age_mean, 0.975)
(ci_age <- c(lowera, uppera))

lowerz <- quantile(samp_199_num$zk_mean, 0.025)
upperz <- quantile(samp_199_num$zk_mean, 0.975)
(ci_zk <- c(lowerz, upperz))


lowery <- quantile(samp_199_num$ye_mean, 0.025)
uppery <- quantile(samp_199_num$ye_mean, 0.975)
(ci_ye <- c(lowery, uppery))

```



Step 10
Finally, use bootstrapping to generate a 95% confidence interval for each variable mean by resampling 1000 samples, with replacement, from your original sample (i.e., by setting the lower and upper CI bounds to 2.5% and 97.5% of the way through the sampling distribution generated by bootstrapping).

```{r}
n_boot <- 10000
boot_height <- vector(length = n_boot)# set up dummy variable FOR height
boot_weight <- vector(length = n_boot)# set up dummy variable FOR weight
boot_age <- vector(length = n_boot)# set up dummy variable FOR age
boot_zk <- vector(length = n_boot)# set up dummy variable FOR zk
boot_ye <- vector(length = n_boot)# set up dummy variable FOR ye
n_size <- 199 #the size of each bootstrap sample should equivalent



##bootstrap for height
for (i in 1:n_boot) {
  boot_height[[i]]<- mean(sample(samp_199_num$height_mean, n_size, replace = T)) #must replace or else we get same vector each time
}


#bootstrap for weight
for (i in 1:n_boot) {
  boot_weight[[i]]<- mean(sample(samp_199_num$weight_mean, n_size, replace = T)) #must replace or else we get same vector each time
}

#boot for age
for (i in 1:n_boot) {
  boot_age[[i]]<- mean(sample(samp_199_num$age_mean, n_size, replace = T)) #must replace or else we get same vector each time
}


#bootstrap for zk
for (i in 1:n_boot) {
  boot_zk[[i]]<- mean(sample(samp_199_num$zk_mean, n_size, replace = T)) #must replace or else we get same vector each time
}

#bootstrap for ye
for (i in 1:n_boot) {
  boot_ye[[i]]<- mean(sample(samp_199_num$ye_mean, n_size, replace = T)) #must replace or else we get same vector each time
}


##cis for bootstrap
lower <- quantile(boot_height, 0.025)
upper <- quantile(boot_height, 0.975)
(ci_boot_height <- c(lower, upper))

lowerww <- quantile(boot_weight, 0.025)
upperww <- quantile(boot_weight, 0.975)
(ci_boot_weight <- c(lowerww, upperww))

loweraa <- quantile(boot_age, 0.025)
upperaa <- quantile(boot_age, 0.975)
(ci_boot_age <- c(loweraa, upperaa))

lowerzz <- quantile(boot_zk, 0.025)
upperzz <- quantile(boot_zk, 0.975)
(ci_boot_zk<- c(lowerzz, upperzz))

loweryy <- quantile(boot_ye, 0.025)
upperyy <- quantile(boot_ye, 0.975)
(ci_boot_ye <- c(loweryy, upperyy))
```
```








